Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates

### 2015-02-17 MS_revised.tex
submitted
ref to figures

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2009}

@@

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf\newline{Article Title - Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates}
}
\newline
% Insert Author names, affiliations and corresponding author email.
\\
Carlos Pe\~na\textsuperscript{1,*},
Marianne Espeland\textsuperscript{2,\textcurrency}
\\
\bf{1} Laboratory of Genetics, Department of Biology, University of Turku, Turku, Finland
\\
\bf{2} Museum of Comparative Zoology and Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, USA
\\

@@

% Current address notes
\textcurrency Current address: McGuire Center for Lepidoptera and Biodiversity,
Florida Museum of Natural History, University of Florida, Gainesville,
FL, US.
% \textcurrency b Insert current address of second author with an address update
% \textcurrency c Insert current address of third author with an address update

@@

\subsection*{Running title: Diversity dynamics in Nymphalidae
butterflies}

@@

\textbf{Keywords:} diversification analysis, MEDUSA, BiSSE, speciation
rate, insect-hostplant dynamics

@@

For analyses, we used the phylogenetic trees from the study of Wahlberg
et al. (2009) \cite{wahlberg2009} that were generated using DNA sequence data from
10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
employed the maximum clade credibility tree (MCC tree) (Fig. 1) as well
as a random sample of 1000 trees from their BEAST run after burnin
\cite{wahlberg2009}. Their original BEAST run was for 40 million generations. We
used a burnin of 25 million generations and took a random sample of 1000
trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (S6 File)
in order to account for phylogenetic uncertainty when performing
the diversification analyses.

@@

Hostplant data for Nymphalidae species were compiled from several
sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
\cite{dyer2002, beccaloni2008, janzen2009} and others (S8--S9 Datasets) for a
total of 6586
hostplant records, including 428 Nymphalidae genera and 143 plant
families and 1070 plant genera. It was not possible to find any
hostplant data for 35 butterfly genera.

@@

We used the statistical software R version 3.0.1 \cite{r2013} in combination
with the APE \cite{popescu2012}, GEIGER \cite{harmon2008}, MEDUSA
\cite{alfaro2009} and
\texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
perform the analyses (included as S10--S12 Supporting Information). All
analyses were run on the 1000 random trees from Wahlberg et al. (2009)
\cite{wahlberg2009}, on the maximum clade credibility tree derived from these and
the MCC tree \cite{wahlberg2009}.

@@

The MEDUSA analysis on the MCC tree in combination with richness data
estimated 18 significant changes in the tempo of diversification in
Nymphalidae history (Fig. 1; Table 1). The corrected AICc acceptance
threshold for adding subsequent piecewise birth-death processes to the
overall model was set to 7.8 units, as prescribed by MEDUSA. In all
MEDUSA analyses, the maximum number of inferred diversification shifts
in all trees was 26. The background net diversification rate for
Nymphalidae was estimated as \texttt{r = 0.092} lineages per Million of
years and the AICc score for the best fit model was \texttt{5090.5}
(Table 1). MEDUSA also estimated that the basic constant birth-death
model was not a better explanation for our data (AICc
\texttt{= 5449.3}).

@@

MultiMEDUSA provided mean and standard deviation statistics for the
diversification values on the shifts on the 1000 trees 
(S16--S17 Text and Figure), and found that the estimated net diversification rate
values had great variation across the posterior distribution of trees. A
boxplot of the net diversification rate values estimated for the clades
that appear in the MCC tree shows that some shifts are estimated as
increased or slowed diversification pace depending on the tree used for
analysis (Fig. 2). This variation is especially wide for the clade
formed by the genera \emph{Magneuptychia} and \emph{Caeruleuptychia}
because MEDUSA estimated diversification values from six times the
background net diversification rate (\texttt{r = 0.5234}) to almost zero
(\texttt{r = 1.22e-01}). The rates for \emph{Taenaris} were between 0.14
and 0.44 (mean value 0.25). Similar degrees of variation were found in
the nodes for \emph{Ypthima}, \emph{Charaxes} and \emph{Coenonympha}
(Fig. 2). The net diversification rates estimates for the clades
(Oleriina + Ithomiina + Napeogenina + Dircennina + Godyridina),
Limenitidinae + Heliconiinae and Pseudergolinae are relatively
consistent across the 1000 trees (Fig. 2).

@@

It is also evident that not all the diversification shifts estimated on
the MCC tree are consistently recovered in most of the 1000 trees. Some
of the shifts in the MCC tree are recovered in very few trees, for
example the shift for the clade Satyrinae is recovered with a
probability of 0.18 (Fig. 3).

@@

The MEDUSA analyses, taking into account phylogenetic uncertainty,
estimated a net diversification rate increase in part of the clade
Ithomiini across more than 95\% of the trees. Our BiSSE analysis found a
positive effect of the character state ``feeding on Solanaceae'' on the
net net diversification rate on part of Ithomiini (Oleriina + Ithomiina
+ Napeogenina + Dircennina + Godyridina) (Fig. 4). The estimated mean
net diversification rate for taxa that do not feed on Solanaceae was
\texttt{r = 0.11} while the net diversification rate for the Solanaceae
feeders was \texttt{r = 0.16} (see S1 Fig. for a boxplot of speciation
and extinction values for the 95\% credibility intervals). The same
analysis considering \emph{Vanessa} and \emph{Hypanartia} as
non-Solanaceae feeders due to dubious records produced the same pattern
and net diversification rates (S2 Fig.). Therefore, the rest of the
analyses were performed assuming these two genera as Solanaceae feeders.
We constrained the BiSSE likelihood model to force equal rates of
speciation for both character states in order to test whether the model
of different speciation rates is a significantly better explanation for
the data. A likelihood ratio test found that the model for increased net
diversification rate for nymphalids feeding on Solanaceae is a
significantly better explanation than this character state having no
effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\)) (Table 3;
character states available in S13 Dataset, code in S18 Supporting Information,
and mcmc run in S19 Supporting Information). We combined the post-burnin mcmc generations
from running BiSSE on 250 trees from the random sample of 1000 trees and
found the same pattern as the BiSSE analysis on the maximum clade
credibility tree (combined mcmc run in S20 Supporting Information; profiles
plot of speciation rates in S3 Fig.; boxplot of 95\% credibility intervals in
S4 Fig.). A BiSSE analysis to test whether the trait ``feeding on
Apocynaceae'' had any effect on increased net diversification rates
found similar speciation rates for lineages feeding on Apocynaceae and
other plants (S5 Fig.).

@@

\begin{figure}[h]
\caption{{\bf BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants}.
Speciation and net diversification
rates are significantly higher in Solanaceae feeders (speciation rate =
$\lambda1$, net diversification rate = $r1$}
\label{fig4}
\end{figure}

@@

The MEDUSA method has been used to infer changes in net diversification
rates in a phylogenetic tree. Since its publication \cite{alfaro2009} the results
of using MEDUSA on a single tree, the maximum clade credibility tree,
have been used for generation of hypotheses and discussion
\cite{heikkila2012, litman2011, ryberg2012}. However, different diversification
shifts and different
rates of diversification are found for certain lineages when
phylogenetic uncertainty was taken into account by using MEDUSA on a
random sample of trees from the posterior distribution of a Bayesian
run. We found that some diversification shifts, estimated on the
Nymphalidae maximum clade credibility tree, were found with very low
probability in the random sample of 1000 trees from the posterior
distribution (Fig. 3, Table 2). We also found that, even though the
analyses estimated the same diversification shifts on two or more trees,
the estimated net diversification rates could vary widely (Fig. 2). For
example, in our Nymphalidae trees, we found that the shift for
\emph{Magneuptychia} and \emph{Caeruleuptychia} had a rate from
\texttt{r = 0.5234}, higher than the background net diversification
rate, to almost zero. This means that observed patterns and conclusions
can be completely contradictory depending on tree choice.

@@

If there is strong phylogenetic signal for increases or decreases in net
diversification rates for a node, it is expected that these shifts would
be inferred by MEDUSA in most of the posterior distribution of trees.
However, weak phylogenetic signal for some nodes can cause some clades
to be absent in some trees and MEDUSA will be unable to estimate any
diversification shift (due to a non-existent node).
The relatively low phylogenetic support for many nodes in the Nymphalidae
tree is likely the reason
why MEDUSA estimated net diversification rate shifts with a probability
higher than 0.90 in the sample of trees for only four shifts: the genus
\emph{Charaxes}, the genus \emph{Ypthima}, part of Ithomiini and the
clade \emph{Callicore} + \emph{Diaethria} (Fig. 3), while estimating
shifts for other lineages with much lower probability.

@@

Our BiSSE analysis, extended to take into account missing taxa and
phylogenetic uncertainty, shows a significantly higher net
diversification rate for Ithomiini taxa, which can be attributed to the
trait ``feeding on Solanaceae hostplants'' (Fig. 4). This is in
agreement with previous findings using other statistical methods
\cite{fordyce2010}. Due to the fact that Ithomiini are virtually the only
nymphalids using Solanaceae as hostplants, it is possible that the trait
responsible for a higher diversification of Ithomiini might not be the
hostplant character. As noted by Maddison et al. (2007) \cite{maddison2007}, the
responsible trait might be a codistributed character such as a trait
related to the ability to digest secondary metabolites.

@@

The increase in diversification rate inferred by MEDUSA occurred after the
probable shift from Apocynaceae to Solanaceae, as the Solanaceae feeders
in the subtribes Melinaeina and Mechanitina are not included in the
diversification shift (shift number 5 in Fig. 1). The apparent
conflicting results from MEDUSA and BiSSE can be explained by the low
species-richness of the subtribes Melinaeina and Mechanitina compared to
the other subtribes included in the shift (52 versus 272 species). It
can be that MEDUSA is more conservative than BiSSE and is not including
Melinaeina and Mechanitina in the shift due to low species numbers.

@@

\section*{Supporting Information Legends}
\subsection*{S1 Fig.}
\label{S1_Fig.}
{\bf Boxplot of speciation (\(\lambda\)) and extinction
(\(\mu\)) values for the 95\% credibility intervals of values estimated
by BiSSE analysis of diversification due to feeding on Solanaceae
plants}.

\subsection*{S2 Fig.}
\label{S2_Fig.}
{\bf BiSSE analysis of diversification of nymphalids
due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
\emph{Hypanartia} as non-Solanaceae feeders. The same pattern is
recovered, speciation and net diversification rates are significantly
higher for Solanaceae feeders \(\lambda1\), r1)}.

\subsection*{S3 Fig.}
\label{S3_Fig.}
{\bf Net diversification rates of nymphalids feeding on
Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
the 1000 trees from the posterior distribution}.

\subsection*{S4 Fig.}
\label{S4_Fig.}
{\bf Boxplot of speciation and extinction values for
the 95\% credibility intervals of values estimated by BiSSE analysis of
diversification due to feeding on Solanaceae plants on the combined
post-burnin runs on 1000 trees from the posterior distribution.}.

\subsection*{S5 Fig.}
\label{S5_Fig.}
{\bf BiSSE analysis of diversification of nymphalids
due to feeding on Apocynaceae hostplants}. Speciation and net
diversification rates are similar.

\subsection*{S6 File}
\label{S6_File}
{\bf 1000 random trees from \cite{wahlberg2009}}.

\subsection*{S7 Dataset}
\label{S7_Dataset}
{\bf Species richness for lineages in Nymphalidae}.

\subsection*{S8Dataset}
\label{S8_Dataset}
{\bf Hostplants of Nymphalidae
butterflies recorded from the literature}.

\subsection*{S9 Dataset}
\label{S9_Dataset}
{\bf References for hostplants data}.

\subsection*{S10 Supporting Information}
\label{S10_Supporting_Information}
{\bf R script to run MEDUSA on the MCC Nymphalidae tree from
\cite{wahlberg2009}}. This script removes the outgroup
taxa and loads the richness data for the tree terminals.

\subsection*{S11 Supporting Information}
\label{S11_Supporting_Information}
{\bf R script to run a MultiMEDUSA analysis on 1000 random trees from
\cite{wahlberg2009}}.

\subsection*{S12 Supporting Information}
\label{S12_Supporting_Information}
{\bf R script to run a MEDUSA
analysis on the MCC tree from the 1000 random trees selected from the
posterior distribution}.

\subsection*{S13 Dataset}
\label{S13_Dataset}
{\bf Data matrix with character states for hosplant use}.

\subsection*{S14 Supporting Information}
\label{S14_Supporting_Information}
{\bf MCC tree from the 1000 random
trees selected from the posterior distribution}.

\subsection*{S15 File}
\label{S15_File}
{\bf Raw results from the MultiMEDUSA run on the random sample of trees from 
the posterior distribution}.

\subsection*{S16 Text}
\label{S16_Text}
{\bf Summary results from a
MultiMEDUSA run on the 1000 random trees selected from the posterior
distribution}.

\subsection*{S17 Figure}
\label{S17_Figure}
{\bf Probability of nodes with
estimated rates from a MultiMEDUSA run on the 1000 random trees selected
from the posterior distribution}.

\subsection*{S18 Supporting Information}
\label{S18_Supporting_Information}
{\bf R script for running the BiSSE analysis}.

\subsection*{S19 Supporting Information}
\label{S19_Supporting_Information}
{\bf Raw results for the BiSSE analysis}.

\subsection*{S20 Supporting Information}
\label{S20_Supporting_Information}
{\bf Raw results for the combined BiSSE analysis}.

\subsection*{S21 Dataset}
\label{S21_Dataset}
{\bf MCC Nymphalidae tree from \cite{wahlberg2009}}

\subsection*{S22 Text}
\label{S22_Text}
{\bf Results of running MEDUSA on the MCC tree}.

\subsection*{S23 Figure}
\label{S23_Figure}
{\bf Summary of the MultiMEDUSA
analysis on the 1000 trees from the posterior distribution}.

\subsection*{S24 Text}
\label{S24_Text}
{\bf Summary results from a MEDUSA
run on the MCC tree from the 1000 random trees selected from the
posterior distribution}.

\subsection*{S25 Figure}
\label{S25_Figure}
{\bf Figure for MEDUSA run on MCC tree from random 1000 trees}.

>>>>

% Use the PLoS provided BiBTeX style
\bibliographystyle{plos2015}

@@

% Title must be 150 characters or less
\begin{flushleft}
{\Large
\textbf\newline{Article Title - Diversity dynamics in Nymphalidae butterflies: Effect of phylogenetic uncertainty on diversification rate shift estimates}
}
\newline
% Insert Author names, affiliations and corresponding author email.
\\
Carlos Pe\~na\textsuperscript{1,*},
Marianne Espeland\textsuperscript{2,3}
\\
\bf{1} Laboratory of Genetics, Department of Biology, University of Turku, Turku, Finland
\\
\bf{2} Museum of Comparative Zoology and Department of Organismic and Evolutionary Biology, Harvard University, Cambridge, USA
\\

@@

% Current address notes
\bf{3} Current address: McGuire Center for Lepidoptera and Biodiversity,
Florida Museum of Natural History, University of Florida, Gainesville,
FL, US.
% \textcurrency b Insert current address of second author with an address update
% \textcurrency c Insert current address of third author with an address update

@@

@@

@@

For analyses, we used the phylogenetic trees from the study of Wahlberg
et al. (2009) \cite{wahlberg2009} that were generated using DNA sequence data from
10 gene regions for 398 of the 540 valid genera in Nymphalidae. We
employed the maximum clade credibility tree (MCC tree) (Fig.~\ref{fig1}) as well
as a random sample of 1000 trees from their BEAST run after burnin
\cite{wahlberg2009}. Their original BEAST run was for 40 million generations. We
used a burnin of 25 million generations and took a random sample of 1000
trees using Burntrees v.0.1.9 \url{http://www.abc.se/~nylander/} (S6 File)
in order to account for phylogenetic uncertainty when performing
the diversification analyses.

@@

Hostplant data for Nymphalidae species were compiled from several
sources including \cite{ackery1988}, HOSTS database (\url{http://bit.ly/YI7nwW}),
\cite{dyer2002, beccaloni2008, janzen2009} and others (S8 and S9 Dataset) for a
total of 6586
hostplant records, including 428 Nymphalidae genera and 143 plant
families and 1070 plant genera. It was not possible to find any
hostplant data for 35 butterfly genera.

@@

We used the statistical software R version 3.0.1 \cite{r2013} in combination
with the APE \cite{popescu2012}, GEIGER \cite{harmon2008}, MEDUSA
\cite{alfaro2009} and
\texttt{diversitree} \cite{fitzjohn2012} packages along with our own scripts to
perform the analyses (included as S10 and S11 File). All
analyses were run on the 1000 random trees from Wahlberg et al. (2009)
\cite{wahlberg2009}, on the maximum clade credibility tree derived from these and
the MCC tree \cite{wahlberg2009}.

@@

The MEDUSA analysis on the MCC tree in combination with richness data
estimated 18 significant changes in the tempo of diversification in
Nymphalidae history (Fig.~\ref{fig1}; Table 1). The corrected AICc acceptance
threshold for adding subsequent piecewise birth-death processes to the
overall model was set to 7.8 units, as prescribed by MEDUSA. In all
MEDUSA analyses, the maximum number of inferred diversification shifts
in all trees was 26. The background net diversification rate for
Nymphalidae was estimated as \texttt{r = 0.092} lineages per Million of
years and the AICc score for the best fit model was \texttt{5090.5}
(Table 1). MEDUSA also estimated that the basic constant birth-death
model was not a better explanation for our data (AICc
\texttt{= 5449.3}).

@@

MultiMEDUSA provided mean and standard deviation statistics for the
diversification values on the shifts on the 1000 trees 
(S16 Text and S17 Figure), and found that the estimated net diversification rate
values had great variation across the posterior distribution of trees. A
boxplot of the net diversification rate values estimated for the clades
that appear in the MCC tree shows that some shifts are estimated as
increased or slowed diversification pace depending on the tree used for
analysis (Fig.~\ref{fig2}). This variation is especially wide for the clade
formed by the genera \emph{Magneuptychia} and \emph{Caeruleuptychia}
because MEDUSA estimated diversification values from six times the
background net diversification rate (\texttt{r = 0.5234}) to almost zero
(\texttt{r = 1.22e-01}). The rates for \emph{Taenaris} were between 0.14
and 0.44 (mean value 0.25). Similar degrees of variation were found in
the nodes for \emph{Ypthima}, \emph{Charaxes} and \emph{Coenonympha}
(Fig.~\ref{fig2}). The net diversification rates estimates for the clades
(Oleriina + Ithomiina + Napeogenina + Dircennina + Godyridina),
Limenitidinae + Heliconiinae and Pseudergolinae are relatively
consistent across the 1000 trees (Fig.~\ref{fig2}).

@@

It is also evident that not all the diversification shifts estimated on
the MCC tree are consistently recovered in most of the 1000 trees. Some
of the shifts in the MCC tree are recovered in very few trees, for
example the shift for the clade Satyrinae is recovered with a
probability of 0.18 (Fig.~\ref{fig3}).

@@

The MEDUSA analyses, taking into account phylogenetic uncertainty,
estimated a net diversification rate increase in part of the clade
Ithomiini across more than 95\% of the trees. Our BiSSE analysis found a
positive effect of the character state ``feeding on Solanaceae'' on the
net net diversification rate on part of Ithomiini (Oleriina + Ithomiina
+ Napeogenina + Dircennina + Godyridina) (Fig.~\ref{fig4}). The estimated mean
net diversification rate for taxa that do not feed on Solanaceae was
\texttt{r = 0.11} while the net diversification rate for the Solanaceae
feeders was \texttt{r = 0.16} (see S1 Figure for a boxplot of speciation
and extinction values for the 95\% credibility intervals). The same
analysis considering \emph{Vanessa} and \emph{Hypanartia} as
non-Solanaceae feeders due to dubious records produced the same pattern
and net diversification rates (S2 Figure). Therefore, the rest of the
analyses were performed assuming these two genera as Solanaceae feeders.
We constrained the BiSSE likelihood model to force equal rates of
speciation for both character states in order to test whether the model
of different speciation rates is a significantly better explanation for
the data. A likelihood ratio test found that the model for increased net
diversification rate for nymphalids feeding on Solanaceae is a
significantly better explanation than this character state having no
effect on diversification (\(\chi^2 = 12.3; 1 df; p < 0.001\)) (Table 3;
character states available in S13 Dataset, code in S18 Supporting Information,
and mcmc run in S19 Supporting Information). We combined the post-burnin mcmc generations
from running BiSSE on 250 trees from the random sample of 1000 trees and
found the same pattern as the BiSSE analysis on the maximum clade
credibility tree (combined mcmc run in S20 Supporting Information; profiles
plot of speciation rates in S3 Figure; boxplot of 95\% credibility intervals in
S4 Figure). A BiSSE analysis to test whether the trait ``feeding on
Apocynaceae'' had any effect on increased net diversification rates
found similar speciation rates for lineages feeding on Apocynaceae and
other plants (S5 Figure).

@@

\begin{figure}[h]
\caption{{\bf BiSSE analysis of diversification of nymphalids due
to feeding on Solanaceae hostplants}.
Speciation and net diversification
rates are significantly higher in Solanaceae feeders (speciation rate =
$\lambda1$, net diversification rate = $r1$)}
\label{fig4}
\end{figure}

@@

The MEDUSA method has been used to infer changes in net diversification
rates in a phylogenetic tree. Since its publication \cite{alfaro2009} the results
of using MEDUSA on a single tree, the maximum clade credibility tree,
have been used for generation of hypotheses and discussion
\cite{heikkila2012, litman2011, ryberg2012}. However, different diversification
shifts and different
rates of diversification are found for certain lineages when
phylogenetic uncertainty was taken into account by using MEDUSA on a
random sample of trees from the posterior distribution of a Bayesian
run. We found that some diversification shifts, estimated on the
Nymphalidae maximum clade credibility tree, were found with very low
probability in the random sample of 1000 trees from the posterior
distribution (Fig.~\ref{fig3}, Table 2). We also found that, even though the
analyses estimated the same diversification shifts on two or more trees,
the estimated net diversification rates could vary widely (Fig.~\ref{fig2}). For
example, in our Nymphalidae trees, we found that the shift for
\emph{Magneuptychia} and \emph{Caeruleuptychia} had a rate from
\texttt{r = 0.5234}, higher than the background net diversification
rate, to almost zero. This means that observed patterns and conclusions
can be completely contradictory depending on tree choice.

@@

If there is strong phylogenetic signal for increases or decreases in net
diversification rates for a node, it is expected that these shifts would
be inferred by MEDUSA in most of the posterior distribution of trees.
However, weak phylogenetic signal for some nodes can cause some clades
to be absent in some trees and MEDUSA will be unable to estimate any
diversification shift (due to a non-existent node).
The relatively low phylogenetic support for many nodes in the Nymphalidae
tree is likely the reason
why MEDUSA estimated net diversification rate shifts with a probability
higher than 0.90 in the sample of trees for only four shifts: the genus
\emph{Charaxes}, the genus \emph{Ypthima}, part of Ithomiini and the
clade \emph{Callicore} + \emph{Diaethria} (Fig.~\ref{fig3}), while estimating
shifts for other lineages with much lower probability.

@@

Our BiSSE analysis, extended to take into account missing taxa and
phylogenetic uncertainty, shows a significantly higher net
diversification rate for Ithomiini taxa, which can be attributed to the
trait ``feeding on Solanaceae hostplants'' (Fig.~\ref{fig4}). This is in
agreement with previous findings using other statistical methods
\cite{fordyce2010}. Due to the fact that Ithomiini are virtually the only
nymphalids using Solanaceae as hostplants, it is possible that the trait
responsible for a higher diversification of Ithomiini might not be the
hostplant character. As noted by Maddison et al. (2007) \cite{maddison2007}, the
responsible trait might be a codistributed character such as a trait
related to the ability to digest secondary metabolites.

@@

The increase in diversification rate inferred by MEDUSA occurred after the
probable shift from Apocynaceae to Solanaceae, as the Solanaceae feeders
in the subtribes Melinaeina and Mechanitina are not included in the
diversification shift (shift number 5 in Fig.~\ref{fig1}). The apparent
conflicting results from MEDUSA and BiSSE can be explained by the low
species-richness of the subtribes Melinaeina and Mechanitina compared to
the other subtribes included in the shift (52 versus 272 species). It
can be that MEDUSA is more conservative than BiSSE and is not including
Melinaeina and Mechanitina in the shift due to low species numbers.

@@

\section*{Supporting Information Legends}
\textbf{S1 Figure.
\bf Boxplot of speciation (\(\lambda\)) and extinction
(\(\mu\)) values for the 95\% credibility intervals of values estimated
by BiSSE analysis of diversification due to feeding on Solanaceae
plants.}

\textbf{S2 Figure.
BiSSE analysis of diversification of nymphalids
due to feeding on Solanaceae hostplants assuming \emph{Vanessa} and
\emph{Hypanartia} as non-Solanaceae feeders}. The same pattern is
recovered, speciation and net diversification rates are significantly
higher for Solanaceae feeders \(\lambda1\), r1).

\textbf{S3 Figure.
Net diversification rates of nymphalids feeding on
Solanaceae plants as estimated by combining post-burnin runs of BiSSE on
the 1000 trees from the posterior distribution}.

\textbf{S4 Figure.
Boxplot of speciation and extinction values for
the 95\% credibility intervals of values estimated by BiSSE analysis of
diversification due to feeding on Solanaceae plants on the combined
post-burnin runs on 1000 trees from the posterior distribution.}

\textbf{S5 Figure.
BiSSE analysis of diversification of nymphalids
due to feeding on Apocynaceae hostplants}. Speciation and net
diversification rates are similar.

\textbf{S6 File.
1000 random trees from \cite{wahlberg2009}}.

\textbf{S7 Dataset.
Species richness for lineages in Nymphalidae}.

\textbf{S8 Dataset.
Hostplants of Nymphalidae
butterflies recorded from the literature}.

\textbf{S9 Dataset.
References for hostplants data}.

\textbf{S10 File.
R script to run MEDUSA on the MCC Nymphalidae tree from
\cite{wahlberg2009}}. This script removes the outgroup
taxa and loads the richness data for the tree terminals.

\textbf{S11 File.
R script to run a MultiMEDUSA analysis on 1000 random trees from
\cite{wahlberg2009}}.

\textbf{S12 Supporting Information.
R script to run a MEDUSA
analysis on the MCC tree from the 1000 random trees selected from the
posterior distribution}.

\textbf{S13 Dataset.
Data matrix with character states for hosplant use}.

\textbf{S14 Supporting Information.
MCC tree from the 1000 random
trees selected from the posterior distribution}.

\textbf{S15 File.
Raw results from the MultiMEDUSA run on the random sample of trees from 
the posterior distribution}.

\textbf{S16 Text.
Summary results from a
MultiMEDUSA run on the 1000 random trees selected from the posterior
distribution}.

\textbf{S17 Figure.
Probability of nodes with
estimated rates from a MultiMEDUSA run on the 1000 random trees selected
from the posterior distribution}.

\textbf{S18 Supporting Information.
R script for running the BiSSE analysis}.

\textbf{S19 Supporting Information.
Raw results for the BiSSE analysis}.

\textbf{S20 Supporting Information.
Raw results for the combined BiSSE analysis}.

\textbf{S21 Dataset.
MCC Nymphalidae tree from \cite{wahlberg2009}}

\textbf{S22 Text.
Results of running MEDUSA on the MCC tree}.

\textbf{S23 Figure.
Summary of the MultiMEDUSA
analysis on the 1000 trees from the posterior distribution}.

\textbf{S24 Text.
Summary results from a MEDUSA
run on the MCC tree from the 1000 random trees selected from the
posterior distribution}.

\textbf{S25 Figure.
Figure for MEDUSA run on MCC tree from random 1000 trees}.
